Prospects of detecting gamma-ray emission from galaxy clusters: cosmic rays and dark matter annihilations

### 2011-04-06 paper.tex
minor changes

%NEW
\newcommand{\rmn}{\mathrm}
\newcommand{\bx}{{\bf x}}
\newcommand{\fg}{{F_\gamma}}
\newcommand{\sg}{{S_\gamma}}
\newcommand{\psf}{\theta_\rmn{res}}
\newcommand{\ph}{\rmn{ph}}
\newcommand{\eph}{E_\ph}
\newcommand{\nph}{N_\ph}
\newcommand{\vir}{\rmn{vir}}
\newcommand{\gal}{\rmn{gal}}
\newcommand{\sd}{\rmn{SD}}
\newcommand{\clu}{\rmn{cl}}
\newcommand{\sfe}{\rmn{sfe}}
\newcommand{\sub}{\rmn{sub}}
\newcommand{\msun}{M_\odot}
\newcommand{\ee}{E_\rmn{e}}
\newcommand{\stars}{\rmn{stars}}
\newcommand{\dust}{\rmn{dust}}
\newcommand{\lx}{L_\rmn{X}}
\newcommand{\s}{\rmn{s}}
\newcommand{\Kp}{\rmn{K}`}
\newcommand{\Ip}{\rmn{I}`}
\newcommand{\Js}{\rmn{J}^*}
\newcommand{\Jp}{\rmn{J}`}
\newcommand{\B}{\rmn{B}}
\newcommand{\bsub}{\B_\rmn{sub}}
\newcommand{\qCR}{q_{\rmn{CR}-\ensuremath{\pi^0}}}

@@

\subsection{Particle physics}
\label{sect:PF}

\subsubsection{Leptophilic models}
\label{sect:LP}
Sommerfeld enhancement, electron spectra

\subsubsection{Supersymmetric dark matter}
neutralino, benchmark models, continuum emission, electron spectra

\subsubsection{Final state radiation}
The photon spectrum from resulting from FSR is universal with only a
weak dependence of the underlying particle physics model. The photon
yield from this process is given by (see
e.g. \cite{2008JHEP...01..049B})
\begin{equation}
\frac{\dd N_{X \bar{X}}}{\dd x} \approx \frac{\alpha Q_X^2}{\pi}
\mathcal{F}_X(x) \log\left[\frac{4 m_\chi^2\left(1-x\right)}{m_X^2}\right]\,.
\end{equation}
Here, the normalized photon energy $x=\eg/m_\chi c^2$, $\alpha =
e^2/\hbar c$ is the fine-structure constant, $Q_X^2$ and $m_X$ the
charge and mass of the particle $X$, respectively. The function
$\mathcal{F}_X(x)$ depends on the spin of the final state and is given
by
\begin{equation}
\mathcal{F}_\rmn{fermion}(x) = \frac{1+\left(1-x\right)^2}{x}\,
\end{equation}
for fermions.

@@

We continue by comparing the estimated differential flux from Fornax
to three other clusters in Fig.~\ref{fig:clu_comp}; the close by well
studied Virgo cluster, the X-ray bright Perseus clusters, and the
massive merging Coma cluster. We find high DM induced gamma-ray fluxes
from both the Fornax and Virgo clusters, which highlight them as
promising targets for indirect DM searches. Note, however, that we
have used spatially unresolved upper limits, which is quite optimistic
for the angular extended Virgo cluster. It is quite striking how
constraining the upper limits for the Fornax cluster is compared to
the other clusters, especially in the $1-10$~GeV energy regime where
Fermi-LAT is most sensitive due to the maximized effective area. The
upper limits for Virgo and Perseus are background contaminated by AGN
activity from M87 and NGC 1267, respectively, and do not gain much
from the increased sensitivity. One should also note that the high CRp
induced flux from both Coma and Perseus is of the same magnitude as
the DM flux from LP models, making it difficult for indirect DM
searches in those clusters. Conversely is the Fornax cluster a great
target for indirect DM studies because the relative low gamma-ray flux
from CRps, absents of an active AGN, and high DM gamma-ray
flux. Actually, the Fermi-LAT 3 year data will be able to constrain
the smallest mass of substructures using the DM BM models. We have
also learned that the closest clusters are superior for DM searches
compared to further away clusters, even though the close by clusters
have a mass that is smaller. In fact, we can understand this by
looking at the luminosity-to-mass scaling relations. The gamma-ray
luminosity from the smooth density profile is given by
\begin{equation}
L_{\gamma,\rmn{sm}} \propto \int \dd V \rho(r)^2 \propto \frac{c^3}
{\left[\log\left(1+c\right)-c/(1+c)\right]} \propto \mvir^{0.83}\,,
\end{equation}
and the total DM luminosity including boost factors for the LP and
the BM models are given by
\begin{eqnarray}
L_{\gamma,\rmn{LP}} &=& L_{\gamma,\rmn{smLP}} \B_\rmn{LP} \propto \mvir^{0.73}\qquad
\,\,\,\,\,\,\rmn{and}\\
L_{\gamma,\rmn{BM}} &=& L_{\gamma,\rmn{smBM}} \B_\rmn{BM} \propto  \mvir^{1.06}\,,\qquad
\rmn{where} \\
B_\rmn{LP}&=&B_\sfe B_\sub \propto \mvir^{-1/3}\mvir^{0.23} \propto
 \mvir^{-0.1}\,,\\
B_\rmn{BM}&=&B_\sub \propto \mvir^{0.23}\,.
\end{eqnarray}
Hence increasing the mass with a factor two doubles the luminosity,
but increasing the distance by a factor two decrease the flux with a
factor four.
\del{
\begin{eqnarray}
L_{\gamma,\rmn{sub tot}} \propto N\,L_{\gamma,\sub} \propto N\,M_\sub 
 \propto M_\sub^{\left(A+1\right)}\,\frac{\dd N_\sub}{\dd M} \propto  M^{A-0.9}
\end{eqnarray}
Assuming that the concentration $c$ decrease for increasing mass as it
does for the mass range $10^{12}-10^{14}\,\msun$ where
$L_{\gamma,\sub} \propto N\,M_\sub^A$, where $A<1$.
}

>>>>

%NEW
\newcommand{\rmn}{\mathrm}
\newcommand{\bx}{{\bf x}}
\newcommand{\fg}{{F_\gamma}}
\newcommand{\sg}{{S_\gamma}}
\newcommand{\psf}{\theta_\rmn{res}}
\newcommand{\ph}{\rmn{ph}}
\newcommand{\eph}{E_\ph}
\newcommand{\nph}{N_\ph}
\newcommand{\vir}{\rmn{vir}}
\newcommand{\gal}{\rmn{gal}}
\newcommand{\sd}{\rmn{SD}}
\newcommand{\clu}{\rmn{cl}}
\newcommand{\sfe}{\rmn{sfe}}
\newcommand{\sub}{\rmn{sub}}
\newcommand{\msun}{M_\odot}
\newcommand{\ee}{E_\rmn{e}}
\newcommand{\stars}{\rmn{stars}}
\newcommand{\dust}{\rmn{dust}}
\newcommand{\lx}{L_\rmn{X}}
\newcommand{\s}{\rmn{s}}
\newcommand{\Kp}{\rmn{K}^\prime}
\newcommand{\Ip}{\rmn{I}^\prime}
\newcommand{\Js}{\rmn{J}^*}
\newcommand{\Jp}{\rmn{J}^\prime}
\newcommand{\B}{\rmn{B}}
\newcommand{\bsub}{\B_\rmn{sub}}
\newcommand{\qCR}{q_{\rmn{CR}-\ensuremath{\pi^0}}}
\newcommand{\ds}{{\sc DarkSUSY}}

@@

\subsection{Detecting Particle Dark Matter}
\label{sect:PF}
Besides the intrinsic interest in the gamma-ray flux from galaxy
clusters generated by conventional hadronic and electromagnetic
process - which should be close to observability with the Fermi-LAT
data \cite{Berezinsky:1996wx,2007A&A...473...41E,2010MNRAS.409..449P}
- the possible contribution from WIMP dark matter is of greatest
interest. A WIMP (Weakly Interactive Massive Particle) which fulfils
the WMAP bounds on the relic density of cold dark matter (CDM) of
\cite{Komatsu:2010fb}
$$\Omega_{CDM}h^2=0.112\pm 0.0056,$$ will in many cases naturally give
a $\gamma$ yield which may be observable. In addition, there are
possible enhancement effects known, such as the astrophysical boost
from dense substructure of dark matter halos, or the particle physics
boost from the Sommerfeld effect, which may increase the chances of
detection further.

There are three methods for detecting dark matter candidates which are
presently employed: First, at particle physics accelerators such as
CERN's LHC, one is now entering into a new energy regime which may
allow the production of the heavy, electrically neutral and long-lived
particles which may constitute dark matter. From these experiments one
may get the first glimpse of the mass scale beyond the standard model
where dark matter may reside. However, to really show that any of the
hypothetical, new particles created at the LHC is actually the dark
matter, one has to rely on the two other methods available for the
dark matter search, namely direct and indirect detection. Direct
detection methods, which are presently evolving rapidly use the feeble
interaction between that dark matter particles of the galactic halo
and nuclei such as Germanium, Sodium, Iodine, Argon or Xenon to infer
the scattering cross section and a rough estimate of the mass of the
dark matter particle (for the currently most sensitive search, see
\cite{Aprile:2010um}). A characteristic of this method is that it
basically only depends on the local dark matter density (which is
rather well determined to be around $0.4$ GeV/cm$^3$) and the basic
cross section for dark matter - nucleus scattering. A disadvantage is
that it does benefit from the two enhancement mechanisms mentioned
above, boost from substructure and/or the Sommerfeld effect. Also, it
cannot be excluded - although it seems at present improbable - that
the local halo structure is such that the solar system happens to be
in an underdense region.

In indirect detection, on the other hand, in particular in the photon
channel, one searches for products of dark matter annihilation in the
galactic halo and beyond. Particularly interesting targets are the
galactic center (for a recent possible indication of a signal, see
\cite{Hooper:2010mq}), the dwarf galaxies surrounding the Milky Way
\cite{Strigari:2006rd,Essig:2009jx,2010JCAP...01..031S}, galaxy
clusters (the focus of this work)
\cite{Ghigna:1998vn,Lewis:2002mfa,Boyarsky:2006kc,2006A&A...455...21C,2009PhRvL.103r1302P},
and even the cosmological large scale structure
\cite{Bergstrom:2001jj,Ullio:2002pj,Taylor:2002zd,Elsaesser:2004ap,Cuoco:2010jb,Abazajian:2010sq,Abdo:2010dk,Zavala:2011tt}.

There is also a possiblity to search for dark matter annihilation
indirectly in antimatter channels, like antiprotons or positrons. Here
diffusion is, however, very efficient which means that the directional
signature which is so important for gamma-rays, is completely
erased. This method has, however, recently been much in focus due to
the surprising rise with energy of the electron to proton ratio
measured by the PAMELA satellite experiment \cite{Adriani:2008zr} and
the ``bump'' in the total electron plus positron spectrum detected by
Fermi \cite{Abdo:2009zk}. The viable particle physics models
describing these results contain some non-trivial, or even unnatural
elements, however. First of all, annhilation to quarks has to be very
much suppressed, due to the non-observation of an enhanced antiproton
flux by PAMELA \cite{Adriani:2010rc}. This means that only so-called
{\em leptophilic} models, where the annihilation only goes to leptons
have to be considered. Here, one has to have a sizeable fraction going
to muons and tau leptons, as the shape of the spectrum disfavours
direct annihilation into electrons and positrons. To get agreement
with the measured absolute rates, one also has to have a large boost
factor (of the order of at least several hundred), which is difficult
to achieve with substrucutre alone. An additional enhancement of the
annihilation rate has to be invoked, for example given by the {\em
  Sommerfeld effect}. This basically means that the usual connection
between annihilation rate at decoupling (when relative velocities $v$
were of order $0.2 - 0.3$), which gives the relic density, and
present-day annihilation in the halo (with velocities $v\sim 10^{-3}$)
is broken. This would come about because the dark matter sector
contains additional light particles which mediates an interaction with
almost bound states which enhance the wave function at the relative
origin, and therefore the annihilation rate (for a pedagogical review
of viable scenarios, see \cite{ArkaniHamed:2008qn}). As a final
difficulty of these models, the lack of an inverse-compton signature
from the galactic center means that only shallow density profiles can
describe the galactic dark matter density distribution
\cite{Bertone:2008xr,Cirelli:2008pk,Bergstrom:2008ag}. (For a review
of the dark matter modelling of these effects, see
\cite{Bergstrom:2009ib}. For a recent treatment, showing still viable
models, see \cite{Finkbeiner:2010sm}.)

A more standard explanation of the PAMELA/Fermi excess would be
pulsars or other supernova remnants (e.g.,
\cite{Hooper:2008kg,Ahlers:2009ae}). In that case, dark matter would
more likely be explained by the standard scenario of a WIMP, for
example the thoroughly studied lightest supersymmertric neutralino
(for reviews, see
\cite{Jungman:1995df,Bergstrom:2000pn,Bertone:2004pz}). We note that
the soon to be launched AMS-02 experiment on the International Space
Station \cite{ams02} may gave interesting clues to the origin of the
antimatter excess.

\subsubsection{Leptophilic models}
\label{sect:LP}
The results of PAMELA \cite{Adriani:2008zr} and Fermi-LAT
\cite{Abdo:2009zk} (and to some extent ATIC
\cite{2008Natur.456..362C}, although the bump seen there at 600 GeV
has not been reproduced by other experiments) has caused intense
activity in the theoretical dark matter community. At first, there was
skepticism whether such a large signal could be compatible with a
WIMP. However, the possibility of an enhanced annihilation rate due to
dark matter halo substructure has been realised for a long time
\cite{1993ApJ...411..439S,Bergstrom:1998zs,Moore:1999nt}, although it
is difficult to achieive the boost factor of the order of a few
hundred to a thousand in the solar neighbourhood, as substructure
according to numerical simulations is most important in the outer
portions of the halo, due to tidal stripping of subhaloes in the inner
part. Another potentially very important effect which may explain the
large boost had been found a few years earlier, namely the Sommerfeld
enhancement. This effect, was computed for electromagnetism by Arnold
Sommerfeld may years ago \cite{sommerfeld} and recently rediscovered
in the quantum field theory of dark matter
\cite{2005PhRvD..71f3528H,2007NuPhB.787..152C,2009PhRvD..79a5014A}.
In quantum mechanics describing electron scattering and electron
positron annihilation, it is caused by the distortion of the plane
wave describing the relative motion of the annihilating particle pair,
due to the near formation of a bound state caused by photon
exchange. In the ladder approximation for QED, one reproduces the
Sommerfeld effect, and the square of the relative wave function at the
origin (which enters into the probablity for the short-distance
process of annihilation) is increased by the factor
\cite{2009PhRvD..79a5014A}
$$
S_{QED}=\frac{|\psi(0)|^2}{|\psi^{(0)}(0)|^2}=|\frac{\frac{\pi\alpha}{v}}{1-e^{-\frac{\pi\alpha}{v}}}|,
$$ with $\alpha =e^2/\hbar c$ the fine-structure constant, and $v$ the
realtive velocity. This amounts to $S_{QED}=\pi\alpha/v$ for small
velocities. For a Yukawa-like particle of mass $m_\phi$, mediating a
weak attractive force with coupling contant $\alpha_Y$ between dark
matter particles $\chi$ with mass $m_\chi$, the small-velocity limit
of the enhancement becomes instead
$$
S_Y\sim\frac{\alpha_Y m_\chi}{m_\phi}.
$$ In the general case, with mediators that may also excite virtual
charged particles in the dark matter sector, one has to solve
numerically a coupled system of differential equations using
appropriate boundary conditions
\cite{2005PhRvD..71f3528H,2007NuPhB.787..152C,2009PhRvD..79a5014A}. In
some cases, the enhancement factor $S$ can be as high as several
hundred to a few thousand, depnding however on the exact parameters of
the theory. The effect is usually strongly velocity-dependent,
depending on velicity as $1/v$ or, in the fine-tuning case of being
very near resonance, as $1/v^2$. This means that in a virialised
system (such a a galaxy cluster) with large velocity dispersion
$v_{\rm cl}$ the Sommerfeld enhancement will be smaller than the one
expected in a single galaxy such as the Milky Way, roughly by a factor
$v_{\rm MW}/v_{\rm cl}$.

As leptophilic models, apart from being slightly contrived from the
particle physics point of view, are also rather limited by several
sets of astrophysical data, we take advantage of the recent reanalysis
\cite{Finkbeiner:2010sm} to define a benchmark leptophilic model that
is still viable.  It is found that the bounds from WMAP5 approximately
imply a Sommerfeld boost factor $S$ which has to satisfy $S(v\to
0)\lesssim 200/f\ (m_Y/1$ GeV$)$, where $f$ is the fraction of energy
from dark matter annihilation that ionizes the intergalactic medium,
$f\sim 0.7$ for annihilation to electrons and a factor a a few smaller
for all other standard model final states.

Taking into account the whole cosmic history of leptophilic models,
also utilizing limit on spectral and polarization distorsions of the
CMB, the maximal value for the boost factor for a 1 to 2 TeV particle
and sub-GeV force carriers is found to be \cite{Finkbeiner:2010sm}
between 400 and 800. Although lower than the first estimates (e.g.,
\cite{Bergstrom:2009fa,Meade:2009iu}) this is still enough to explain
the PAMELA and Fermi excess, given other astrophysical uncertainties.

To connect easily to our previous work \cite{2009PhRvL.103r1302P}, we
choose as our benchmark leptophilic model one with mass $1.6$ TeV and
Sommerfeld boost $S=300$.

\subsubsection{Supersymetric dark matter}
The most studied models for dark matter are supersymmetric ones,
especially models where the lightest supersymmetric particle, in most
models the lightest neutralino, is stable. The stability is assured in
viable SUSY models due to a discrete symmetry, $R$-parity. This
symmetry is needed from the particle physics point of view to avoid
fast proton decay, and automatically makes the lightest supersymmetric
particle stable, and therefore a good candidate for dark matter. In
addition, as the coupling to ordinary matter is given by gauge
couplings, the neutralino is automatically a WIMP
candidate. Unfortunately, the breaking of supersymmetry (which has to
be there since otherwise, e.g., there would exist a scalar electron
with the same mass as the electron) means that a large number of
essentially free parameters enter actual calculations.  The neutralino
is then a linear combination of the supersymmetric partners of the
U(1) gauge boson, one of the SU(2) gauge bosons, and the two Higgs
doublets in the Minimal Supersymmetric extension of the Standard Model
(the MSSM) with mixing coefficients which depend on the SUSY breaking
parameters \cite{1984NuPhB.238..453E}. The parameter space is very
large, too large in general to scan efficiently even with the most
powerful computers. Therefore, often simplifying assumptions are made
to limit the number of free parameters. Even then, a large number of
viable models is found which give a relic density consistent with the
WMAP data. This means that they are indeed very good templates for
WIMP dark matter, something that explains their popularity besides the
subjective statement that few more plausible models for dark matter
have been put forward so far.

The the wave function of the lightest neutralino can be written
$$
\tilde\chi_1=a_1\tilde B+a_2\tilde W_3+a_3 \tilde H_1^0+a_4\tilde H_2^0,
$$
with 
$$
\sum_{i=1}^4 |a_i|^2=1.
$$
the gaugino fraction of a given neutralino is $|a_1|^2+|a_2|^2$ whereas $|a_3|^2+|a_4|^2$ is the higgsino fraction.

We use the \ds package \cite{ds} to compute the mass, couplings and
relic density for a given set of parameters.  Actually, we take
advantage of so-called benchmark models, which have been proposed in
different contexts (in particular in \cite{Bringmann:2008kj} giving
predictions for the MAGIC experiment), and which give a good
representation of models with high enough gamma-ray rates to become
the first candidates for dark matter indirect detection in imaging air
Cherenkov telescopes. We use the following supersymmetric benchmark
models:

\begin{itemize}
\item
 $I′$: This model was introduced in \cite{2004EPJC...33..273B}, where
  its phenomenology at colliders was studied. Annihilation directly
  into lepton pairs is suppressed for neutralinos due to their
  Majorana nature and therefore helicity suppression for annihilation
  in the galactic halo \cite{1983PhRvL..50.1419G}. The higher order
  process (inner bremsstrahlung, IB) $\tilde\chi_1\tilde\chi_1\to
  \ell^+\ell^-\gamma$, which does not suffer from helicity suppression
  \cite{1989PhLB..225..372B,2008JHEP...01..049B}, gives a considerable
  contribution due to light sleptons in this model.

\item $J′$: This model, also from \cite{2004EPJC...33..273B} is in the
  so-called coannihilation tail. The sleptons are nearly degenerate
  with the neutralino, which causes the large IB from the leptonic
  final states to give a high enhancement of the flux.

\item $K′$: A representative model for the funnel region, where the
  annihilation dominantly occurs in the $s$-channel through exchange
  of the pseudoscalar Higgs boson \cite{2004EPJC...33..273B}. Here IB
  contributions are not important.

\item $J^*$: Annihilation in the so-called coannihilation region,
  introduced in \cite{2008JHEP...01..049B} as BM3, with a particularly
  large IB contribution.

\end{itemize}

\subsubsection{Final state radiation}
There are two types of radiative processes which are important for our
benchmark models, first the usual QED final state radiation (FSR) for
both the leptophilic model and for supersymmetric models with charged
final states. In addition, there may be direct emission from a
virtual, charged, exchanged particle (such as a spin-0 slepton or
squark), internal bremsstrahlung, IB. The latter is essentially only
important when there is a helicity suppression for the lowest order
annihilation process, which is the case for neutralinos, whose
Majorana nature make the annihilation rate in the $s$ wave
proportional to $m_f^2$, for final state fermion $f$. Interestingly,
both the gamma-ray and fermion energy spectra are peaked at the
highest energy for this process, which can in some models cause a
rather spectacular bump in the spectra for these models.

The photon spectrum resulting from FSR is universal with only a
weak dependence of the underlying particle physics model. The photon
yield from this process is given by (see
e.g. \cite{2008JHEP...01..049B})
\begin{equation}
\frac{\dd N_{X \bar{X}}}{\dd x} \approx \frac{\alpha Q_X^2}{\pi}
\mathcal{F}_X(x) \log\left[\frac{4 m_\chi^2\left(1-x\right)}{m_X^2}\right]\,.
\end{equation}
Here, the normalized photon energy $x=\eg/m_\chi c^2$, 
$\alpha =e^2/\hbar c$ is the fine-structure constant, 
$Q_X^2$ and $m_X$ the
charge and mass of the particle $X$, respectively. The function
$\mathcal{F}_X(x)$ depends on the spin of the final state and is given
by
\begin{equation}
\mathcal{F}_\rmn{fermion}(x) = \frac{1+\left(1-x\right)^2}{x}\,
\end{equation}
for fermions.

The differential energy spectrum for the IB process is more
complicated (see \cite{1989PhLB..225..372B,2008JHEP...01..049B}), but
can be computed using {\sc DarkSUSY}. The differential photon energy
spectra for all our benchmark models are shown in
Fig.~\ref{fig:bmspectra}.  The peaking at high photon energy and the
resulting bump caused by inner bremsstrahlung is clearly seen for
benchmark models $I´$, $J´$, and $J^*$.

@@

We continue by comparing the estimated differential flux from Fornax
to three other clusters in Fig.~\ref{fig:clu_comp}; the close by well
studied Virgo cluster, the X-ray bright Perseus clusters, and the
massive merging Coma cluster. We find high DM induced gamma-ray fluxes
from both the Fornax and Virgo clusters, which highlight them as
promising targets for indirect DM searches. Note, however, that we
have used spatially unresolved upper limits, which is quite optimistic
for the angular extended Virgo cluster. It is quite striking how
constraining the upper limits for the Fornax cluster is compared to
the other clusters, especially in the $1-10$~GeV energy regime where
Fermi-LAT is most sensitive due to the maximized effective area. The
upper limits for Virgo and Perseus are background contaminated by AGN
activity from M87 and NGC 1267, respectively, and do not gain much
from the increased sensitivity. One should also note that the high CRp
induced flux from both Coma and Perseus is of the same magnitude as
the DM flux from LP models, making it difficult for indirect DM
searches in those clusters. Conversely is the Fornax cluster a great
target for indirect DM studies because the relative low gamma-ray flux
from CRps, absents of an active AGN, and high DM gamma-ray
flux. Actually, the Fermi-LAT 3 year data will be able to constrain
the smallest mass of substructures using the DM BM models. We have
also learned that the closest clusters are superior for DM searches
compared to further away clusters, even though the close by clusters
have a mass that is smaller. In fact, we can understand this by
looking at the luminosity-to-mass scaling relations. The gamma-ray
luminosity from the smooth density profile is given by
\begin{equation}
L_{\gamma,\rmn{sm}} \propto \int \dd V \rho(r)^2 \propto \frac{c^3}
{\left[\log\left(1+c\right)-c/(1+c)\right]} \propto \mvir^{0.83}\,,
\end{equation}
and the total DM luminosity including boost factors for the LP and
the BM models are given by
\begin{eqnarray}
L_{\gamma,\rmn{LP}} &=& L_{\gamma,\rmn{smLP}} \B_\rmn{LP} \propto \mvir^{0.73}\qquad
\,\,\,\,\,\,\rmn{and}\\
L_{\gamma,\rmn{BM}} &=& L_{\gamma,\rmn{smBM}} \B_\rmn{BM} \propto  \mvir^{1.06}\,,\qquad
\rmn{where} \\
B_\rmn{LP}&=&B_\sfe B_\sub \propto \mvir^{-1/3}\mvir^{0.23} \propto
 \mvir^{-0.1}\,,\\
B_\rmn{BM}&=&B_\sub \propto \mvir^{0.23}\,.
\end{eqnarray}
Hence increasing the mass with a factor two doubles the luminosity,
but increasing the distance by a factor two decrease the flux with a
factor four. It is also usefull to relate the luminosities to a more
model independent variable as the maximum circular velocity given by
\begin{eqnarray}
V_\rmn{max} &=& V_\rmn{circ}(r_\rmn{max})\qquad\,\,\,\,\,\,\,\,\,\rmn{where}\nonumber\\
V_\rmn{circ}(r)&=& \sqrt{\frac{G_\rmn{N}\,M(<r)}{r}}\qquad\rmn{and}\nonumber\\
\frac{\dd V_\rmn{circ}(r)}{\dd r} &=& 0\quad \Longrightarrow\quad r_\rmn{max}\,.
\end{eqnarray}
We find that $V_\rmn{max}=650\,\rmn{km/s}$ of a large $10^{15}\,\msun$
cluster, and $V_\rmn{max}=420\,\rmn{km/s}$ of a small
$4\times10^{13}\,\msun$ group.

\del{
\begin{eqnarray}
L_{\gamma,\rmn{sub tot}} \propto N\,L_{\gamma,\sub} \propto N\,M_\sub 
 \propto M_\sub^{\left(A+1\right)}\,\frac{\dd N_\sub}{\dd M} \propto  M^{A-0.9}
\end{eqnarray}
Assuming that the concentration $c$ decrease for increasing mass as it
does for the mass range $10^{12}-10^{14}\,\msun$ where
$L_{\gamma,\sub} \propto N\,M_\sub^A$, where $A<1$.
}
